% !TEX root = ../../main.tex

% \chapter{Evaluation}
% \label{ch:eval}

This last part of the thesis presents an empirical evaluation of \system.
We first demonstrate the implementation of a realistic collaborative application
atop the middleware.
With this as our main benchmark, we then evaluate the platform
experimentally, comparing it to a classical client-server approach in
the cloud, and to a simple caching approach.
We consider both the online and the offline case.
In the former, we evaluate transaction throughput vs.~response time, and
behavior under load.
In the offline case, we measure reconnection
time, i.e., the time it takes for disconnected clients to be
synchronized again.
We also evaluate the performance benefit of peer groups.
Finally, we study migration in mobile setups, measuring
the time to return to normal performance.

\begin{notice}
    \system{}, the examples used in the evaluation, the scripts to launch
    the evaluation and the scripts to parse the results are all available at the following
    url: \url{https://gitlab.inria.fr/itoumlilt/colony}
\end{notice}


\chapter{\system{}Chat benchmark application}
\label{ch:eval:benchmark-app}
\input{benchmark}

\chapter{Performance Evaluation}

The performance evaluation will focus on situations where all clients are 
continuously online, 
as well as on situations where the network is interrupted. 
For online situations, 
we are especially interested in the time it takes to distribute and apply an 
update to all other clients that are editing the same data. 
For the offline situation, 
we are especially interested in the time it takes for all clients to get back 
in sync with each other after the network disruption, 
and in the time it takes to restore normal interactive performance.

\section{Setup}
\label{sec:experimental-setup}
\input{setup}

\section{Response time and throughput}
\label{sec:eval1}
\input{eval1}

\section{Response time of offline collaboration}
\label{sec:eval2}
\input{eval2}

\section{Migration effect on response time}
\label{sec:eval3}
\input{eval3}

\section{Performance compared to related work}
\label{sec:eval4}
\input{eval4}

\chapter{Summary}
\input{discussion}
